---
title: "ST 558 Project 2"
author: "Lee Bennett"
format: html
editor: visual
---

```{r}
#| include: false

library(tidyverse)
library(readxl)
library(geofacet)
```

## Introduction

In this notebook, we'll do some static exploratory analysis of the US superstore sales data (<https://www.kaggle.com/datasets/juhi1994/superstore/data>) before creating a dashboard in R Shiny for interactive analysis. The data was downloaded as the file "US Superstore data.xls" and we'll load it into R as a tibble:

```{r, warning=FALSE}

sales_raw<-read_excel("US Superstore data.xls") |> as_tibble() |> rename_with(make.names) |> mutate(OrderYear=year(Order.Date))
```

This processed data has been saved as an RData file called `superstore_data.rds` for use in the app.

## Data Exploration

We'll start the exploratory analysis by creating a series of contingency tables, beginning with one-way tables for the region and segment of sales, followed by a two-way table for those same variables:

```{r}
tab1 <- table(sales_raw$Region)
tab1

tab2 <- table(sales_raw$Segment)
tab2

tab3 <- table(sales_raw$Region,sales_raw$Segment)
tab3
```

Next, let's summarize the gross sales and profit variables by region and segment:

```{r}
sales_raw |> group_by(Region) |> summarize(across(c(Sales,Profit),list("mean"=mean,"median"=median,"sd"=sd),.names="{.fn}_{.col}"))

sales_raw |> group_by(Segment) |> summarize(across(c(Sales,Profit),list("mean"=mean,"median"=median,"sd"=sd),.names="{.fn}_{.col}"))
```
Finally, we'll create some plots to explore this data:

```{r}
stat_s <- sales_raw |> group_by(Region,State,Segment,OrderYear) |> summarize(across(c(Sales,Profit),list("sum"=sum, "mean"=mean,"median"=median,"sd"=sd),.names="{.fn}_{.col}")) 

#Total sales by region and segment
graph_1 <- ggplot(stat_s, aes(x=Region, y=sum_Sales, fill=Segment)) + geom_bar(position="dodge",stat="identity") + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Region", y="Total sales ($)", title="Total Sales by Region and Market Segment")

#Total profit by region and segment
graph_2 <- ggplot(stat_s, aes(x=Region, y=sum_Profit, fill=Segment)) + geom_bar(position="dodge",stat="identity") + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Region", y="Total profit ($)", title="Total Profit by Region and Market Segment")

#Sales by Year and Region,
graph_3 <- ggplot(stat_s, aes(x=Region, y=sum_Sales, fill=Segment)) + geom_bar(position="dodge",stat="identity") + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Region", y="Total sales ($)", title="Total Sales by Region, Market Segment, and Year") + facet_wrap(~OrderYear)

graph_4 <- ggplot(sales_raw, aes(x=OrderYear, y=Sales, rank, fill = Segment)) +
  geom_bar(stat="identity") + facet_geo(~ State) + theme_bw()

print(graph_1)
print(graph_2)
print(graph_3)
print(graph_4)
```